---
title: "Cortical-Subcortical Structural Connectivity Supports TMS Engagement of the Amygdala"
author: "Valerie Jill Sydnor"
output:
  html_document:
    theme: journal
    toc: yes
    toc_depth: 4
    toc_float: no
    code_folding: show
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(dplyr)
library(knitr)
library(ppcor)
library(correlation)
library(lsr)
```

&nbsp;  


**Read in Data**

*Participant Demographics and TMS Administration Information*
```{r}
demographics <- read.csv("../../output_measures/ZAPR01_DWI_TMSfMRI_Demographics.csv", header = TRUE)
```

*Tract-Based Mean Fixel Measures*
```{r}
fixels <- read.csv("../../output_measures/fixel_measures/AmygdalaSOS-LHAmygdalaROI25-FixelMeasures.csv", header = TRUE)
```

*Single Pulse TMS-fMRI BOLD Signal Change (Evoked Response) Measures*
```{r}
#Amygdala-targeted vlPFC TMS
vlPFC.TMS.ER.HO25 <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreAmygdala_TMSfMRI_EvokedResponse_HO25_Subcortical.csv", header=FALSE, sep='\t')
vlPFC.TMS.ER.HO25 <- vlPFC.TMS.ER.HO25[,-(3:4)]
colnames(vlPFC.TMS.ER.HO25) <- c("Case","File","left_wm","left_gm","left_ventricle","left_thalamus","left_caudate","left_putamen","left_pallidum","brainstem","left_hippocampus","left_amygdala","left_accumbens","right_wm","right_gm","right_ventricle","right_thalamus","right_caudate","right_putamen","right_pallidum","right_hippocampus","right_amygdala","right_accumbens")

vlPFC.TMS.ER.HO50 <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreAmygdala_TMSfMRI_EvokedResponse_HO50_Subcortical.csv", header=FALSE, sep='\t')
vlPFC.TMS.ER.HO50 <- vlPFC.TMS.ER.HO50[,-(3:4)]
colnames(vlPFC.TMS.ER.HO50) <- c("Case","File","left_wm","left_gm","left_ventricle","left_thalamus","left_caudate","left_putamen","left_pallidum","brainstem","left_hippocampus","left_amygdala","left_accumbens","right_wm","right_gm","right_ventricle","right_thalamus","right_caudate","right_putamen","right_pallidum","right_hippocampus","right_amygdala","right_accumbens")

vlPFC.TMS.ER.sgROI <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreAmygdala_TMSfMRI_EvokedResponse_sgACC.csv",header=FALSE, sep='\t')
vlPFC.TMS.ER.sgROI <- vlPFC.TMS.ER.sgROI[,-(3:4)]
colnames(vlPFC.TMS.ER.sgROI) <- c("Case","File","sgACC")

vlPFC.TMS.ER.aparcaseg <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreAmygdala_TMSfMRI_EvokedResponse_FreeSurfer-LeftAmygdala.csv", header=FALSE, sep='\t')
vlPFC.TMS.ER.aparcaseg <- vlPFC.TMS.ER.aparcaseg[,-(3:4)]
colnames(vlPFC.TMS.ER.aparcaseg) <- c("Case","File","left_amygdala")

#sgACC-target dmPFC TMS
dmPFC.TMS.ER.HO25 <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreSgACC_TMSfMRI_EvokedResponse_HO25_Subcortical.csv", header=FALSE, sep='\t')
dmPFC.TMS.ER.HO25 <- dmPFC.TMS.ER.HO25[,-(3:4)]
colnames(dmPFC.TMS.ER.HO25) <- c("Case","File","left_wm","left_gm","left_ventricle","left_thalamus","left_caudate","left_putamen","left_pallidum","brainstem","left_hippocampus","left_amygdala","left_accumbens","right_wm","right_gm","right_ventricle","right_thalamus","right_caudate","right_putamen","right_pallidum","right_hippocampus","right_amygdala","right_accumbens")

dmPFC.TMS.ER.HO50 <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreSgACC_TMSfMRI_EvokedResponse_HO50_Subcortical.csv", header=FALSE, sep='\t')
dmPFC.TMS.ER.HO50 <- dmPFC.TMS.ER.HO50[,-(3:4)]
colnames(dmPFC.TMS.ER.HO50) <- c("Case","File","left_wm","left_gm","left_ventricle","left_thalamus","left_caudate","left_putamen","left_pallidum","brainstem","left_hippocampus","left_amygdala","left_accumbens","right_wm","right_gm","right_ventricle","right_thalamus","right_caudate","right_putamen","right_pallidum","right_hippocampus","right_amygdala","right_accumbens")

dmPFC.TMS.ER.sgROI <- read.csv("../../output_measures/TMSfMRI_evokedresponse/PreSgACC_TMSfMRI_EvokedResponse_sgACC.csv",header=FALSE, sep='\t')
dmPFC.TMS.ER.sgROI <- dmPFC.TMS.ER.sgROI[,-(3:4)]
colnames(dmPFC.TMS.ER.sgROI) <- c("Case","File","sgACC")
```

*Amygdala-targeted TMS Site of Stimulation MNI Coordinates*
```{r}
SOS.coordinates <- read.csv("../../output_measures/functional_connectivity/PreAmygdala_Target_Coordinates.csv", header=TRUE)
```


## TMS to the ventrolateral prefrontal cortex modulates fMRI BOLD activity in the amgydala 
&nbsp;  

**Single pulses of TMS administered to amygdala functional connectivity peaks located within (or nearest to) the left vlPFC elict a functional response in the amygdala**

```{r, echo=F, eval=T}
print("Mean evoked response HO25")
mean(abs(vlPFC.TMS.ER.HO25$left_amygdala))
print("SD evoked response HO25")
sd(abs(vlPFC.TMS.ER.HO25$left_amygdala))
print("Mean evoked response HO50")
mean(abs(vlPFC.TMS.ER.HO50$left_amygdala))
print("SD evoked response HO50")
sd(abs(vlPFC.TMS.ER.HO50$left_amygdala))
```

**Left amygdala TMS evoked responses are nearly equivalent when defined by individual Freesurfer parcellations**

```{r}
result <- cor.test(vlPFC.TMS.ER.HO25$left_amygdala, vlPFC.TMS.ER.aparcaseg$left_amygdala, method=c("pearson"))
result
cor_to_ci(cor = result$estimate, n = 45, ci = 0.95, method = "pearson")
```


**BOLD activity in the left amygdala decreased in response to vlPFC TMS pulses in the majority of individuals**

```{r}
sum(vlPFC.TMS.ER.HO25$left_amygdala < 0)
t.test(vlPFC.TMS.ER.HO25$left_amygdala, alternative = "two.sided") 
```

**The magnitude of the left amygdala TMS evoked response significantly correlated with the effective strength of neurostimulation, i.e., with *TMS dose***

```{r}
demographics[38,55] <- mean(na.omit(demographics$dose_weighted_vlPFC)) # impute single missing weighted dose by averaging all other doses
result <- cor.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), demographics$dose_weighted_vlPFC, method=c("spearman"), exact = F)
result
cor_to_ci(cor = result$estimate, n = 45, ci = 0.95, method = "spearman")
```

**The magnitude of the left amygdala TMS evoked response did not correlate with absolute stimulator output**

```{r}
result <- cor.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), demographics$stimulator_output_vlPFC, method=c("spearman"), exact=F)
result
cor_to_ci(cor = result$estimate, n = 45, ci = 0.95, method = "spearman")
```

**ER magnitude in the left amygdala did not significantly correlate with ER magnitude in most other subcortical structures, revealing that subcortical responses to TMS are unique and differentiable**

```{r}
vlPFC.TMS.ER.HO25.leftsubcortical <- vlPFC.TMS.ER.HO25[,c(6,7,8,9,11,13)] #extract lh subcortical data

cortests.ER.rho <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(cor.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), method=c("spearman"), exact=F))$estimate})) #Rho from a correlation of magnitude of evoked response in left amygdala and magnitude of evoked response in other subcortical structures
cortests.ER.pval <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(cor.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), method=c("spearman"), exact=F))$p.value})) #p-value from a correlation of magnitude of evoked response in left amygdala and magnitude of evoked response in other subcortical structures
cortests.ER.pval.fdr <- p.adjust(cortests.ER.pval[1,], method = c("fdr")) #FDR-corrected p-value from the paired t-tests

vlPFC.TMS.ERMagnitude.Correlations <- rbind(cortests.ER.rho, cortests.ER.pval, cortests.ER.pval.fdr) #merge measures into one data frame
vlPFC.TMS.ERMagnitude.Correlations <- t(round(vlPFC.TMS.ERMagnitude.Correlations, 6)) #transpose to view results in a table
colnames <- c("rho","p.value","FDR.p.value") #label column data
colnames(vlPFC.TMS.ERMagnitude.Correlations) <- colnames
vlPFC.TMS.ERMagnitude.Correlations
```

**Evoked responses in the left amygdala are almost universally larger than responses in other subcortical structures**

```{r}
means <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){ mean(abs(x))})) #mean magnitude evoked response for each lh subcortical region

t.tests.ER.tstat <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(t.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), paired=T))$statistic})) #t-statistic from a paired t-test comparing evoked responses in the left amygdala and every other lh subcortical region
t.tests.ER.pval <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(t.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), paired=T))$p.value})) #p-value from a paired t-test comparing evoked responses in the left amygdala and every other lh subcortical region
t.tests.ER.pval.fdr <- p.adjust(t.tests.ER.pval[1,], method = c("fdr")) #FDR-corrected p-value from the paired t-tests
t.tests.ER.CIlower <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(t.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), paired=T))$conf.int[1]})) #95% confidence interval from a paired t-test comparing evoked responses in the left amygdala and every other lh subcortical region
t.tests.ER.CIupper <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(t.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), paired=T))$conf.int[2]})) #95% confidence interval from a paired t-test comparing evoked responses in the left amygdala and every other lh subcortical region
t.tests.ER.CohensD <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x){(cohensD(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(x), method="paired"))})) #cohens D from a paired t-test comparing evoked responses in the left amygdala and every other lh subcortical region

vlPFC.TMS.ERMagnitude.Ttests <- rbind(means, t.tests.ER.tstat, t.tests.ER.pval, t.tests.ER.pval.fdr, t.tests.ER.CohensD, t.tests.ER.CIlower, t.tests.ER.CIupper) #merge measures into one data frame
vlPFC.TMS.ERMagnitude.Ttests <- t(round(vlPFC.TMS.ERMagnitude.Ttests, 6)) #transpose to view results in a table
colnames <- c("Mean.ER","t.statistic","p.value","FDR.p.value", "Cohens d", "95% CI lower", "95% CI upper") #label column data
colnames(vlPFC.TMS.ERMagnitude.Ttests) <- colnames
vlPFC.TMS.ERMagnitude.Ttests
```


## Pathway fiber density is associated with the magnitude of the TMS evoked amygdala response 
&nbsp;  

**Higher fiber density in an identified vlPFC-amygdala white matter pathway was associated with a larger TMS-evoked change in left amygdala BOLD signal, controlling for age**

```{r}
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), demographics$age, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

**vlPFC-amygdala pathway cross section was not associated with the TMS-evoked response in the left amygdala, controlling for age and estimated total intracranial volume**

```{r}
covs <- data.frame(demographics$age, demographics$eTIV)
result <- pcor.test(fixels$logFC_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

**Sensitivity Analyses**

* FD-evoked response association controlling for age and estimated total intracranial volume (eTIV, from FreeSurfer stats files)
```{r}
covs <- data.frame(demographics$age, demographics$eTIV)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and sex
```{r}
covs <- data.frame(demographics$age, demographics$gender)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and TMS dose
```{r}
covs <- data.frame(demographics$age, demographics$dose_weighted_vlPFC)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and site of stimulation-amygdala functional connectivity (baseline fMRI scan)
```{r}
covs <- data.frame(demographics$age, demographics$rs.b.GSRDNS.amitLBLA)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and TMS site of stimulation Y and Z MNI coordinates
```{r}
covs <- data.frame(demographics$age, SOS.coordinates$Y, SOS.coordinates$Z)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and DWI scan motion (mean framewise displacement)
```{r}
covs <- data.frame(demographics$age, demographics$DWI_motion_meanFD)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association controlling for age and TMS-fMRI scan motion (rel mean RMS)
```{r}
covs <- data.frame(demographics$age, demographics$PreAmygdala_TMSfMRI_RelMeanRMS)
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_amygdala), covs, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association using Harvard Oxford amygdala ROI with a stricter anatomical threshold (maxprob-thr50)
```{r}
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO50$left_amygdala), demographics$age, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association using individual-specific left amygdala parcellations generated with Freesurfer 
```{r}
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.aparcaseg$left_amygdala), demographics$age, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

* FD-evoked response association at varying fixel streamline mapping thresholds (tck2fixel thresholds)
```{r}
fixel.FD.thresholds <- fixels[,c(2,3,5,6,7)]
cor.tests.threshold.rho <- as.data.frame(lapply(fixel.FD.thresholds, function(x) {pcor.test(x,abs(vlPFC.TMS.ER.HO25$left_amygdala), demographics$age, method=c("spearman"))$estimate})) #rho of FD-evoked response partial correlation at varying thresholds
cor.tests.threshold.p.value <- as.data.frame(lapply(fixel.FD.thresholds, function(x) {pcor.test(x,abs(vlPFC.TMS.ER.HO25$left_amygdala), demographics$age, method=c("spearman"))$p.value})) #p-value of FD-evoked response partial correlation at varying thresholds
FD.ER.Threshold.Cortests <- t(rbind(cor.tests.threshold.rho,cor.tests.threshold.p.value)) #merge and transpose to view results 
colnames <- c("rho","p.value")
colnames(FD.ER.Threshold.Cortests) <- colnames
FD.ER.Threshold.Cortests
```


## The identified pathway is differentially associated with neurostimulation-induced subcortical responses
&nbsp;  

**Higher fiber density in the vlPFC-amygdala pathway was additionally associated with a larger TMS evoked response in the left hippocampus, controlling for age**

```{r}
result <- pcor.test(fixels$FD_mean_5, abs(vlPFC.TMS.ER.HO25$left_hippocampus), demographics$age, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```

**However, fiber density in the vlPFC-amygdala pathway was not associated with TMS evoked response magnitude in any other left hemisphere subcortical structure**

```{r}
cor.tests.lhsubcortical.rho <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x) {pcor.test(abs(x),fixels$FD_mean_5, demographics$age, method=c("spearman"))$estimate})) #rho of FD-evoked response partial correlation for remaining lh subcortical regions
cor.tests.lhsubcortical.pval <- as.data.frame(lapply(vlPFC.TMS.ER.HO25.leftsubcortical, function(x) {pcor.test(abs(x),fixels$FD_mean_5, demographics$age, method=c("spearman"))$p.value})) #p-value of FD-evoked response partial correlation for remaining lh subcortical regions
cor.tests.lhsubcortical.pval.fdr <- p.adjust(cor.tests.lhsubcortical.pval[1,], method=c("fdr"))
FD.ER.LHSubcortical.Cortests <- t(rbind(cor.tests.lhsubcortical.rho,cor.tests.lhsubcortical.pval,cor.tests.lhsubcortical.pval.fdr))
colnames <- c("rho","p.value", "FDR.p.value")
colnames(FD.ER.LHSubcortical.Cortests) <- colnames
FD.ER.LHSubcortical.Cortests
```


## Pathway fiber density is not related to amygdala TMS evoked response when stimulating a distant control site
&nbsp;  

*Left amygdala evoked response characteristics when targeting dorsal/posterior control sites (chosen based on sgACC functional connectivity)*
```{r, echo=F, eval=T}
print("Mean evoked response HO25")
mean(abs(dmPFC.TMS.ER.HO25$left_amygdala))
print("SD evoked response HO25")
sd(abs(dmPFC.TMS.ER.HO25$left_amygdala))
print("Number of individuals with a negative ER")
sum(dmPFC.TMS.ER.HO25$left_amygdala < 0)
```


**The magnitude of left amygdala evoked responses tends to be larger when targeting the vlPFC site (personalized based on amygdala functional connectivity) than when targeting the control site (personalized based on sgACC functional connectivity), though not significantly so**

```{r}
amygdala.ER.difference <- (abs(vlPFC.TMS.ER.HO25$left_amygdala)-abs(dmPFC.TMS.ER.HO25$left_amygdala)) # amygdala-targeting sites ER magnitude - sgACC-targeting sites ER magnitude
(sum(amygdala.ER.difference > 0))/45 # percent of individuals with a larger amygdala ER when targeting the vlPFC
```

```{r}
shapiro.test(amygdala.ER.difference) # test to see whether the differences in ER are normally distributed: p-value = 2.236e-07
wilcox.test(abs(vlPFC.TMS.ER.HO25$left_amygdala), abs(dmPFC.TMS.ER.HO25$left_amygdala), paired=TRUE, mu=0, conf.int = T) #wilcoxin signed rank test that the null distribution of x-y is symmetric about mu =0
```


**When targeting the spatially distant control site, there was no association between vlPFC-amygdala pathway fiber density and the magnitude of the TMS-evoked response in the left amygdala**

```{r}
result <- pcor.test(fixels$FD_mean_5, abs(dmPFC.TMS.ER.HO25$left_amygdala), demographics$age, method=c("spearman"))
result
cor_to_ci(cor = result$estimate, n = result$n, ci = 0.95, method = "spearman")
```
